Robust calibration of numerical models based on relative regret

### 2020-01-20 article.tex
modifs Élise, et template JCP

%% For including figures, graphicx.sty has been loaded in
%% elsarticle.cls. If you prefer to use the old commands
%% please give \usepackage{epsfig}

%% The amssymb package provides various useful mathematical symbols
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
%% The amsthm package provides extended theorem environments
%% \usepackage{amsthm}

%% The lineno packages adds line numbers. Start line numbering with
%% \begin{linenumbers}, end it with \end{linenumbers}. Or switch it on
%% for the whole article with \linenumbers.
%% \usepackage{lineno}
\usepackage{easy-todo}
\usepackage{graphicx}
\usepackage{booktabs}
% À changer
\newcommand{\yobs}{\mathbf{y}^o}
\DeclareMathOperator*{\argmin}{arg\,min \,}
\DeclareMathOperator*{\argmax}{arg\,max \,}
\newcommand{\Var}{\mathbb{V}\textrm{ar}}
\newcommand{\Ex}{\mathbb{E}}
\newcommand{\Prob}{\mathbb{P}}
\newcommand{\Cov}{\textsf{Cov}}
\newcommand{\tra}{\mathrm{tr}}
\newcommand{\kmean}{{\mathbf{k}}_{\Ex}}
\newcommand{\hatkmean}{\hat{\mathbf{k}}_{\Ex}}
\newcommand{\kvar}{{\mathbf{k}}_{\mathbb{V}}}
\newcommand{\hatkvar}{\hat{\mathbf{k}}_{\mathbb{V}}}
\newcommand{\kmpe}{{\mathbf{k}}_{\mathrm{MPE}}}
\newcommand{\hatkmpe}{\hat{\mathbf{k}}_{\mathrm{MPE}}}
\newcommand{\kest}{\hat{\mathbf{k}}}
\newcommand{\RRE}{RRE}
\newcommand{\checkap}{{\alpha}_p}
\newcommand{\checka}{{\alpha}}
\newcommand{\checkk}{\mathbf{k}}
\newcommand{\checkkp}{{\mathbf{k}}_p}
\newcommand{\hatkp}{\hat{\mathbf{k}}_p}
\newcommand{\Kspace}{\mathbb{K}}
\newcommand{\Uspace}{\mathbb{U}}
\newcommand{\JBH}{J_{\mathrm{BH}}}
\newcommand{\JBHS}{J_{\mathrm{BHswap}}}

\newcommand{\arthur}[1]{{\itshape\color{cyan} ({#1})}}
\newcommand{\elise}[1]{{\itshape\color{red} ({#1})}}
\newcommand{\victor}[1]{{\itshape\color{green} ({#1})}}

@@

\title{Robust calibration of numerical models based on relative regret}

%% use optional labels to link authors explicitly to addresses:
%% \author[label1,label2]{}
%% \address[label1]{}
%% \address[label2]{}

\author[vt]{Victor Trappler}
\ead{victor.trappler@univ-grenoble-alpes.fr}
\author[ea]{Élise Arnaud}
\author[av]{Arthur Vidard}
\author[ld]{Laurent Debreu}

\address[vt]{Univ. Grenoble Alpes, CNRS, Inria, Grenoble INP*, LJK, 38000 Grenoble, France}

@@

\begin{keyword}
%% keywords here, in the form: keyword \sep keyword
Calibration \sep Robust optimisation \sep Relative-regret \sep shallow-water equations
%% PACS codes here, in the form: \PACS code \sep code

%% MSC codes here, in the form: \MSC code \sep code
%% or \MSC[2008] code \sep code (2000 is the default)

\end{keyword}

@@

%% main text
\section*{Introduction}
\label{sec:intro}

%----------------------------------------------------------------------------
% \listoftodos{}

%\arthur{Revoir les notations ($\hat{\bf k}$, ${\bf k}^*$ c'est la meme chose ou pas ? les u, U je ne suis pas sûr que c'est bon non plus)}
%\victor{Pour moi, ce qui est en majuscule, c'est des v.a., et tous les inputs sont en gras (mais on peut laisser tomber le gras)}
%% \introduction[modified heading if necessary]
%\subsection{Motivation}

Numerical models are widely used to study or forecast natural phenomena and improve industrial processes. However, by essence models only partially represent reality and sources of uncertainties are ubiquitous (discretisation errors, missing physical processes, poorly known boundary conditions).
%Many physical phenomena are modelled numerically in order to better understand and/or to predict their behaviour. Because of the complexity and or ignorance of the underlying physics, numerical models are imperfect representations of the reality. 
%
%Various sources of uncertainties can corrupt the model quality, such as the discretisation of the domain, or presence of  small scale phenomena that can not be fully represented. 
Moreover, such uncertainties may be of different nature. 
\cite{walker_defining_2003} proposes to consider two categories of uncertainties. On the one hand aleatoric uncertainties,  coming from the inherent variability of a phenomenon, e.g.\ intrinsic randomness of some environmental variables. On the other hand, epistemic uncertainties coming from a lack of knowledge about the properties and conditions of the phenomenon underlying the behaviour of the system under study.
The latter can be accounted for through the introduction of ad-hoc correcting terms in the numerical model, that need to be properly estimated. Thus, reducing the epistemic uncertainty can be done through parameters estimation approaches. 
%when one is having observations of the reality. 
This is usually done using optimal control techniques, leading to an optimisation of a well chosen cost function which is typically built as a comparison with reference observations.
%
An application of such an approach, in the context of ocean circulation modeling, is the estimation of ocean bottom friction parameters in~\cite{das_estimation_1991} and~\cite{boutet_estimation_2015}.

@@

As mentioned before, robustness can be understood as satisfactory for a broad range of $\mathbf{u}$, and/or as insensitive as possible to uncertainties encompassed in $\mathbf{U}$. Under this definition, one may design robust estimators, either by using the moments of the cost function; or by exploiting the distribution of its minimisers.

%
%When robustness is defined as the ability of $\kest$ to perform reasonably well under different operating conditions (i.e.\ different $\mathbf{u}$), two classical  approaches can be used to define robust estimators:
%\begin{itemize} 
%\item Optimisation of the statistical moments of the random variable $J(\mathbf{k}, \mathbf{U})$: usually minimisation of the expected value, of the variance or a combination of them
%\item Estimate based on the distribution of the minimisers $\mathbf{k}_{\mathbf{u}}$, assuming the inverse problem is well-posed to ensure the existence and the unicity of the minimiser for each $\mathbf{u}$. 
%\end{itemize}

%\subsection{simple/classical estimators \arthur{TITRE \`A REVOIR}}

%Broadly speaking, two different aspects can be used to define robust estimators:
%\begin{itemize} 
%\item Optimisation of the statistical moments of the random variable: usually minimisation of the expected value, of the variance or a combination of them
%\item Estimate based on the distribution of the minimisers, assuming the inverse problem is well-posed to ensure the existence and the unicity of the minimiser for each $\mathbf{u}$. 
%\end{itemize}

% ----------------------------------------------------------------------------
\subsection{Optimisation of the moments}

%Statistical moments, especially  expectation and variance, are common tools used to describe and summarise a random variable. %A case in point are the Gaussian random variables, that are uniquely defined by their expected value and their variance. 
%It then makes sense that the statistical moments have a central role in robust optimisation. In practice, such ideas relies on the expected value as seen in \cite{shapiro_lectures_2009}, while looking at the same time mean and variance is the basis of Markowitz's portfolio selection theory in \cite{markowitz_portfolio_1952}. 
%

@@

In general, an analytical form of $p_{\mathbf{K}^*}$ is usually impossible to obtain, so an estimation $\hat{p}_{\mathbf{K}^*}$ must be used, and its maximum computed to get the MPE.
In the rest of the paper, the hat notation will indicate an estimation using numerical values of the underlying theoretical quantity.
Once again, a set of samples  $\{\mathbf{u}^i\}_{i=1\dots N}$ can be used to compute the set $\{\mathbf{k}_{\mathbf{u}^i}\}_{i=1\dots N}$, from which one can 
approximate  $p_{\mathbf{K}^*}$. The resulting approximation and therefore its mode, is sensitive to the density estimation method. Main methods are KDE  (Kernel Density Estimation)~\cite{elise}, and EM (Expectation-Maximisation)~\citep{dempster_maximum_1977}.

@@

\begin{figure*}[!ht]
  \centering
  \includegraphics[width=10cm]{Figures/branin_side_moments_noscale.pdf}
  \caption{The left column concerns $\JBH$, while the right one deals with $\JBHS$. Contours of both functions are plotted on the top, and curves of  $\mu(\mathbf{k})$ and $\sigma(\mathbf{k})$ are shown on the bottom (respective scales are not displayed). Estimates $\hatkmean$ and $\hatkvar$ are plotted with the dashed line.}
\label{fig:branin_moments}
\end{figure*}
%\clearpage % Pour lisibilité dans le preview

@@

Similarly, estimations of $\hatkmpe$ are depicted on Figure~\ref{fig:contours}.The top row shows the contours of both functions as well as the set of conditional minimisers $\{\mathbf{k}_{\mathbf{u}^i}\}_{1\leq i \leq N}$ in red, as defined in Eq.~\eqref{eq:conditional_minimiser}.
The bottom row presents three approximations of the density of minimisers: the histogram in grey (bin size selected using Freedman-Diaconis from~\cite{freedman_histogram_1981}), the result of a kernel density estimation (KDE) with Gaussian kernels in red (using Scott's rule from~\cite{scott_optimal_1979} for bandwidth selection), and the estimation by a Gaussian mixture, calculated with the EM algorithm. The number of Gaussians has been fixed to 3, a guess based on the general shape of the histogram. Respective estimations of $\hatkmpe$ are also depicted using dashed lines. 
\begin{figure}[!h]
  \centering
  \includegraphics[width=10cm]{Figures/branin_side_66.pdf}
  \caption{Top Left:  $\JBH$ along with conditional minimisers in red. Bottom left: Estimated densities using KDE and EM algorithm and the histogram. The dashed lines indicate the MPE found using those methods. Right: Same quantities with $\JBHS$.}
\label{fig:contours}
\end{figure}

@@

\section{Relative regret-based family of estimators}%A new family of robust estimators \elise{titre a revoir}}
\label{sec:relax_constraint}
\subsection{Relaxing the optimality constraint}
The density of minimisers has been estimated by optimising $\mathbf{k}\mapsto J(\mathbf{k},\mathbf{u})$ over $\Kspace$ for different realisations of $\mathbf{u}$. Instead of focusing on optimal values, we propose to consider their {\it acceptable} neighbourhood in terms of performance of the cost function as well.
To do so, for each $\mathbf{u}$, $\mathbf{k}$ is deemed acceptable when  $J(\mathbf{k},\mathbf{u}) \leq \alpha J^*(\mathbf{u})$, with $\alpha > 1$.

In this context, for a given $\mathbf{k}$, the set $R_{\alpha}(\mathbf{k}) \subset \Uspace$ is defined as the set of $\mathbf{u}$, for which $\mathbf{k}$ is acceptable:
\begin{equation}
\label{eq:def_Ralpha}
R_{\alpha}(\mathbf{k}) = \left\{ \mathbf{u} \in \Uspace \mid J(\mathbf{k},\mathbf{u}) \leq \alpha J^*(\mathbf{u}) \right\}
\end{equation}
Figure~\ref{fig:relax_tuto} details the successive steps for the construction of the set $R_{\alpha}$. First, the conditional minimisers are computed, as shown on the top plots. Afterwards, for a given level $\alpha=1.5$ and for all $\mathbf{u} \in \Uspace$, the acceptable $\mathbf{k}$ are identified, as shown on the bottom left plot. Finally, the region $R_{\alpha}(\mathbf{k})$ is the set of $\mathbf{u}$ for which $\mathbf{k}$ is acceptable, as represented with a vertical slice on the bottom right plot.

@@

\begin{figure*}[!t]
\centering
\includegraphics[width=10cm]{Figures/relaxation_tuto.pdf}
\caption{Principle of the relaxation of the constraint on $\JBH$, and illustration of $R_\alpha(\mathbf{k})$. Top plots: Computation of the conditional minimisers $\mathbf{k}^*(\mathbf{u})$. Top right plot: the set $(\mathbf{k}^*(\mathbf{u}),\mathbf{u})$ of conditional minimisers is represented in red. On the bottom left plot, for a relaxation $\alpha=1.5$, and $\mathbf{u}=3$, the acceptable $\mathbf{k}$ are in cyan, while the non-acceptable ones are in orange, while the frontier $\{(\mathbf{k}^*(\mathbf{u}),\mathbf{u}) \mid J^*(\mathbf{u}) = J(\mathbf{k},\mathbf{u})\}$ is in yellow. On the bottom right plot, the set $R_{\alpha}(\mathbf{k})$ for $\alpha=1.5$ and $\mathbf{k}=1.5$ is in green.}
\label{fig:relax_tuto}
\end{figure*}

@@

Noting that without relaxation, \textit{i.e.} when $\alpha$ is set to $1$, $\Gamma_1$ is non-zero if the set $\{\mathbf{u}\in\Uspace \mid J(\mathbf{k},\mathbf{u}) = J^*(\mathbf{u})\}$ has non-zero measure with respect to $\Prob_{\mathbf{U}}$. It happens when the distribution of $\mathbf{K}^*$ presents atoms.

This can be linked to the definition of the distribution of the minimisers $\mathbf{K}^*$. For instance, if $\Kspace$ is a discrete set, we can rewrite $\Gamma_1$ as $\Gamma_1(\mathbf{k})=\Prob_{\mathbf{U}}\left[J(\mathbf{k},\mathbf{u}) = J^*(\mathbf{U})\right] = \Prob_{\mathbf{U}}\left[\mathbf{k} = \mathbf{k}^*(\mathbf{U})\right]$ which is the probability mass function of $\mathbf{K}^*$.

The motivation behind this relaxation is to take into account the local behaviour of the function around the conditional minimisers.
For a given set of environmental conditions $\mathbf{u}$, if the function $\mathbf{k} \mapsto J(\mathbf{k},\mathbf{u})$ is flat around its minimum $\mathbf{k}^*(\mathbf{u})$, then choosing  $\mathbf{k}^*(\mathbf{u}) + \epsilon$ (for a small $\epsilon$) will produce a value closer to the minimum than when the function has a high curvature.
In addition to that, relaxing the constraint using a multiplicative constant puts more weight on the values of $\mathbf{k}^*(\mathbf{u})$ when $J^*(\mathbf{u})$ is small.

The choice of the relaxation constant $\alpha$ can be made to ensure the existence of a parameter that is ``acceptable'' with a certain probability.
For instance, given that $J>0$, $\Gamma_{\alpha}(\mathbf{k})$ is increasing with respect to $\alpha$ for any $\mathbf{k}\in\Kspace$. We can then focus on the smallest value of $\alpha$ such that $\Gamma_\alpha$ reaches a certain level of confidence $p\in[0,1]$. This leads to the definition of $\checkap$
\begin{align}
  \checkap &= \inf\left\{ \alpha\geq 1 \mid \exists \checkkp \in \Kspace,\, \Gamma_{\alpha}(\checkkp) \geq p \right\} \nonumber \\
   &= \inf \left\{ \alpha \geq 1 \mid \max_{\mathbf{k}\in\Kspace} \Gamma_{\alpha}(\mathbf{k}) \geq p \right\}   \label{eq:def_alpha_check}
\end{align}
that is the smallest $\alpha$, such that there exists a particular $\checkkp \in \Kspace$ for which $J(\checkkp,\mathbf{U}) \leq \checkap J^*(\mathbf{U})$ with probability $p$.
$\{\mathbf{k}_p \text{ for } p \in [0;1] \}$ defines the relative-regret family of robust estimators (\RRE).
This formulation can be linked to the quantiles and Value-at-Risk of the random variable $\max_{\mathbf{k}} \{ \frac{J(\mathbf{k},\mathbf{U})}{J^*(\mathbf{U})}\}$, that is a measure of risk usually applied in the financial sector (see~\cite{rockafellar_deviation_2002}). For different levels $p$, and thus different $\alpha_p$, Figure~\ref{fig:illu_alpha_p} shows examples of $\Gamma_{\alpha_p}$. By definition, the associated $\mathbf{k}_p$ is then the first value for which $\Gamma_{\alpha_p}$ reaches $p$. We can see that changing the level $p$ shifts the maximiser of $\Gamma_{\alpha_p}$, and that for small $\alpha$, $\Gamma_{\alpha}(\mathbf{k}_1)$ is also very small. This indicates that $\mathbf{k}_1$ is located quite far from the conditional minimisers, and arise as a compromise when the relaxation is large enough.

\begin{figure}[!ht]
\centering
\includegraphics[width = 0.6\textwidth]{Figures/illu_alpha_p.pdf}
\caption{Illustration of the influence of different levels $p$ on $\Gamma_{\alpha_p}$ and on $\mathbf{k}_p$}
\label{fig:illu_alpha_p}
\end{figure}

@@

By considering the particular case where $p=1$, a relation between $\checkk_1$ and $\checka_1$ can be derived, by using Eq.~\eqref{eq:def_alpha_check} and the strict positivity of the cost function:
\begin{equation}
  \label{eq:def_alpha1}
  \Prob_{\mathbf{U}}\left[\frac{J(\mathbf{k}_1,\mathbf{U})}{ J^*(\mathbf{U}) } \leq \checka_1\right]=1
\end{equation}
it follows then that 
\begin{equation}
  \label{eq:upper_bound_alpha_check}
  \checka_1 = \sup_{\mathbf{u}\in\Uspace} \frac{J(\mathbf{k}_1,\mathbf{u})}{J^*(\mathbf{u})} =  \inf_{\mathbf{k}\in\Kspace} \left\{ \sup_{\mathbf{u}\in\Uspace} \frac{J(\mathbf{k},\mathbf{u})}{J^*(\mathbf{u})} \right\} 
\end{equation}

>>>>

%% For including figures, graphicx.sty has been loaded in
%% elsarticle.cls. If you prefer to use the old commands
%% please give \usepackage{epsfig}
\usepackage{jcomp}
\usepackage{framed,multirow}
%% The amssymb package provides various useful mathematical symbols
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
%% The amsthm package provides extended theorem environments
%% \usepackage{amsthm}

%% The lineno packages adds line numbers. Start line numbering with
%% \begin{linenumbers}, end it with \end{linenumbers}. Or switch it on
%% for the whole article with \linenumbers.
%% \usepackage{lineno}
\usepackage{graphicx}
\usepackage{booktabs}
% À changer
\newcommand{\yobs}{\mathbf{y}^o}
\DeclareMathOperator*{\argmin}{arg\,min \,}
\DeclareMathOperator*{\argmax}{arg\,max \,}
\newcommand{\Var}{\mathbb{V}\textrm{ar}}
\newcommand{\Ex}{\mathbb{E}}
\newcommand{\Prob}{\mathbb{P}}
\newcommand{\Cov}{\textsf{Cov}}
\newcommand{\tra}{\mathrm{tr}}
\newcommand{\kmean}{{\mathbf{k}}_{\Ex}}
\newcommand{\hatkmean}{\hat{\mathbf{k}}_{\Ex}}
\newcommand{\kvar}{{\mathbf{k}}_{\mathbb{V}}}
\newcommand{\hatkvar}{\hat{\mathbf{k}}_{\mathbb{V}}}
\newcommand{\kmpe}{{\mathbf{k}}_{\mathrm{MPE}}}
\newcommand{\hatkmpe}{\hat{\mathbf{k}}_{\mathrm{MPE}}}
\newcommand{\kest}{\hat{\mathbf{k}}}
\newcommand{\RRE}{RRE}
\newcommand{\checkap}{{\alpha}_p}
\newcommand{\checka}{{\alpha}}
\newcommand{\checkk}{\mathbf{k}}
\newcommand{\checkkp}{{\mathbf{k}}_p}
\newcommand{\hatkp}{\hat{\mathbf{k}}_p}
\newcommand{\Kspace}{\mathbb{K}}
\newcommand{\Uspace}{\mathbb{U}}
\newcommand{\JBH}{J_{\mathrm{BH}}}
\newcommand{\JBHS}{J_{\mathrm{BHswap}}}

\newcommand{\arthur}[1]{{\itshape\color{cyan} ({#1})}}
\newcommand{\elise}[1]{{\itshape\color{red} ({#1})}}
\newcommand{\victor}[1]{{\itshape\color{green} ({#1})}}
\usepackage{url}
\usepackage{xcolor}
\definecolor{newcolor}{rgb}{.8,.349,.1}

@@

% \title{Robust calibration of numerical models based on relative regret}

% %% use optional labels to link authors explicitly to addresses:
% %% \author[label1,label2]{}
% %% \address[label1]{}
% %% \address[label2]{}

% \author[vt]{Victor Trappler}
% \ead{victor.trappler@univ-grenoble-alpes.fr}
% \author[ea]{Élise Arnaud}
% \author[av]{Arthur Vidard}
% \author[ld]{Laurent Debreu}

% \address[vt]{Univ. Grenoble Alpes, CNRS, Inria, Grenoble INP*, LJK, 38000 Grenoble, France}

\verso{Victor Trappler \textit{etal}}

\title{Robust calibration of numerical models based on relative regret}%

\author[1]{Victor \snm{Trappler}\corref{cor1}}
\cortext[cor1]{Corresponding author% : 
  % Tel.: +0-000-000-0000;  
  % fax: +0-000-000-0000
  % ;
}
\ead{victor.trappler@univ-grenoble-alpes.fr}
\author[1]{Élise \snm{Arnaud}}
% \fntext[fn1]{This is author footnote for second author.}  
\author[2]{Arthur \snm{Vidard}}
%% Third author's email
% \ead{author3@author.com}
\author[2]{Laurent \snm{Debreu}}

\address[1]{Univ. Grenoble Alpes, CNRS, Inria, Grenoble INP\fnref{fn1}, LJK, 38000 Grenoble, France}
\fntext[fn1]{Institute of Engineering Univ. Grenoble Alpes}
\address[2]{Univ. Grenoble Alpes, Inria, CNRS, Grenoble INP\fnref{fn1}, LJK, 38000 Grenoble, France}
\received{}
\finalform{}
\accepted{}
\availableonline{}
\communicated{}

@@

% \begin{keyword}
% %% keywords here, in the form: keyword \sep keyword
% Calibration \sep Robust optimisation \sep Relative-regret \sep shallow-water equations
% %% PACS codes here, in the form: \PACS code \sep code

% %% MSC codes here, in the form: \MSC code \sep code
% %% or \MSC[2008] code \sep code (2000 is the default)

% \end{keyword}

@@

%% main text
\section{Introduction}
\label{sec:intro}

%----------------------------------------------------------------------------
Numerical models are widely used to study or forecast natural phenomena and improve industrial processes. However, by essence models only partially represent reality and sources of uncertainties are ubiquitous (discretisation errors, missing physical processes, poorly known boundary conditions).
%Many physical phenomena are modelled numerically in order to better understand and/or to predict their behaviour. Because of the complexity and or ignorance of the underlying physics, numerical models are imperfect representations of the reality. 
%
%Various sources of uncertainties can corrupt the model quality, such as the discretisation of the domain, or presence of  small scale phenomena that can not be fully represented. 
Moreover, such uncertainties may be of different nature. 
\cite{walker_defining_2003} proposes to consider two categories of uncertainties. On the one hand aleatoric uncertainties,  coming from the inherent variability of a phenomenon, e.g.\ intrinsic randomness of some environmental variables. On the other hand, epistemic uncertainties coming from a lack of knowledge about the properties and conditions of the phenomenon underlying the behaviour of the system under study.
The latter can be accounted for through the introduction of ad-hoc correcting terms in the numerical model, that need to be properly estimated. Thus, reducing the epistemic uncertainty can be done through parameters estimation approaches. 
%when one is having observations of the reality. 
This is usually done using optimal control techniques, leading to an optimisation of a well chosen cost function which is typically built as a comparison with reference observations.
%
An application of such an approach, in the context of ocean circulation modeling, is the estimation of ocean bottom friction parameters in~\cite{das_estimation_1991} and~\cite{boutet_estimation_2015}.

@@

As mentioned before, robustness can be understood as satisfactory for a broad range of $\mathbf{u}$, and/or as insensitive as possible to uncertainties encompassed in $\mathbf{U}$. Under this definition, one may design robust estimators, either by using the moments of the cost function; or by exploiting the distribution of its minimisers.
% ----------------------------------------------------------------------------
\subsection{Optimisation of the moments}

@@

In general, an analytical form of $p_{\mathbf{K}^*}$ is usually impossible to obtain, so an estimation $\hat{p}_{\mathbf{K}^*}$ must be used, and its maximum computed to get the MPE.
In the rest of the paper, the hat notation will indicate an estimation using numerical values of the underlying theoretical quantity.
Once again, a set of samples  $\{\mathbf{u}^i\}_{i=1\dots N}$ can be used to compute the set $\{\mathbf{k}_{\mathbf{u}^i}\}_{i=1\dots N}$, from which one can 
approximate  $p_{\mathbf{K}^*}$. The resulting approximation and therefore its mode, is sensitive to the density estimation method. Main methods are KDE (Kernel Density Estimation)~\citep{silverman_density_2018}, and EM (Expectation-Maximisation)~\citep{dempster_maximum_1977}.

@@

\begin{figure*}[!ht]
  \centering
  \includegraphics[width=12cm]{Figures/branin_side_moments_noscale.pdf}
  \caption{The left column concerns $\JBH$, while the right one deals with $\JBHS$. Contours of both functions are plotted on the top, and curves of  $\mu(\mathbf{k})$ and $\sigma(\mathbf{k})$ are shown on the bottom (respective scales are not displayed). Estimates $\hatkmean$ and $\hatkvar$ are plotted with the dashed line.}
\label{fig:branin_moments}
\end{figure*}
%\clearpage % Pour lisibilité dans le preview

@@

Similarly, estimations of $\hatkmpe$ are depicted on Figure~\ref{fig:contours}. The top row shows the contours of both functions as well as the set of conditional minimisers $\{\mathbf{k}_{\mathbf{u}^i}\}_{1\leq i \leq N}$ in red, as defined in Eq.~\eqref{eq:conditional_minimiser}.
The bottom row presents three approximations of the density of minimisers: the histogram in grey (bin size selected using Freedman-Diaconis from~\cite{freedman_histogram_1981}), the result of a kernel density estimation (KDE) with Gaussian kernels in red (using Scott's rule from~\cite{scott_optimal_1979} for bandwidth selection), and the estimation by a Gaussian mixture, calculated with the EM algorithm. The number of Gaussians has been fixed to 3, a guess based on the general shape of the histogram. Respective estimations of $\hatkmpe$ are also depicted using dashed lines. 
\begin{figure}[!h]
  \centering
  \includegraphics[width=12cm]{Figures/branin_side_66.pdf}
  \caption{Top Left:  $\JBH$ along with conditional minimisers in red. Bottom left: Estimated densities using KDE and EM algorithm and the histogram. The dashed lines indicate the MPE found using those methods. Right: Same quantities with $\JBHS$.}
\label{fig:contours}
\end{figure}

@@

\section{Relative regret-based family of estimators}%A new family of robust estimators \elise{titre a revoir}}
\label{sec:relax_constraint}
\subsection{Relaxing the optimality constraint}
The density of minimisers has been estimated by optimising $\mathbf{k}\mapsto J(\mathbf{k},\mathbf{U}=\mathbf{u})$ over $\Kspace$ for different realisations of $\mathbf{u}$. Instead of focusing on optimal values, we propose to consider their {\it acceptable} neighbourhood in terms of performance of the cost function as well.
To do so, for each $\mathbf{u}$, $\mathbf{k}$ is deemed acceptable when  $J(\mathbf{k},\mathbf{U}=\mathbf{u}) \leq \alpha J^*(\mathbf{u})$, with $\alpha > 1$. 
%
%
In this context, for a given $\mathbf{k}$, the set $R_{\alpha}(\mathbf{k}) \subseteq \Uspace$ is defined as the set of $\mathbf{u}$, for which $\mathbf{k}$ is acceptable:
\begin{equation}
\label{eq:def_Ralpha}
R_{\alpha}(\mathbf{k}) = \left\{ \mathbf{u} \in \Uspace \mid J(\mathbf{k},\mathbf{U}=\mathbf{u}) \leq \alpha J^*(\mathbf{u}) \right\}
\end{equation}
Figure~\ref{fig:relax_tuto} details the successive steps for the construction of the set $R_{\alpha}$. First, the conditional minimisers are computed, as shown on the top plots. Afterwards, for a given level $\alpha=1.5$, the set of acceptable $\mathbf{k}$ can be identified for each $\mathbf{u} \in \Uspace$, as shown on the bottom left plot. Finally, the region $R_{\alpha}(\mathbf{k})$ is the subset of $\Uspace$ for which $\mathbf{k}$ is acceptable, as represented with a vertical slice on the bottom right plot.

@@

\begin{figure*}[!t]
\centering
\includegraphics[width=12cm]{Figures/relaxation_tuto.pdf}
\caption{Principle of the relaxation of the constraint on $\JBH$, and illustration of $R_\alpha(\mathbf{k})$. Top plots: Computation of the conditional minimisers $\mathbf{k}^*(\mathbf{u})$. Top right plot: the set $(\mathbf{k}^*(\mathbf{u}),\mathbf{u})$ of conditional minimisers is represented in red. On the bottom left plot, for a relaxation $\alpha=1.5$, and $\mathbf{u}=3$, the acceptable $\mathbf{k}$ are in cyan, while the non-acceptable ones are in orange, while the frontier $\{(\mathbf{k}^*(\mathbf{u}),\mathbf{u}) \mid J^*(\mathbf{u}) = J(\mathbf{k},\mathbf{U}=\mathbf{u})\}$ is in yellow. On the bottom right plot, the set $R_{\alpha}(\mathbf{k})$ for $\alpha=1.5$ and $\mathbf{k}=1.5$ is in green.}
\label{fig:relax_tuto}
\end{figure*}

@@

Noting that without relaxation, \textit{i.e.} when $\alpha$ is set to $1$, $\Gamma_1$ is non-zero if the set $\{\mathbf{u}\in\Uspace \mid J(\mathbf{k},\mathbf{U}=\mathbf{u}) = J^*(\mathbf{u})\}$ has non-zero measure with respect to $\Prob_{\mathbf{U}}$. It happens when the distribution of $\mathbf{K}^*$ presents atoms.

This can be linked to the definition of the distribution of the minimisers $\mathbf{K}^*$. For instance, if $\Kspace$ is a discrete set, we can rewrite $\Gamma_1$ as $\Gamma_1(\mathbf{k})=\Prob_{\mathbf{U}}\left[J(\mathbf{k},\mathbf{U}=\mathbf{u}) = J^*(\mathbf{U})\right] = \Prob_{\mathbf{U}}\left[\mathbf{k} = \mathbf{k}^*(\mathbf{U})\right]$ which is the probability mass function of $\mathbf{K}^*$.

The motivation behind this relaxation is to take into account the local behaviour of the function around the conditional minimisers.
For a given set of environmental conditions $\mathbf{u}$, if the function $\mathbf{k} \mapsto J(\mathbf{k},\mathbf{U}=\mathbf{u})$ is flat around its minimum $\mathbf{k}^*(\mathbf{u})$, then choosing  $\mathbf{k}^*(\mathbf{u}) + \epsilon$ (for a small $\epsilon$) will produce a value closer to the minimum than when the function has a high curvature.
In addition to that, relaxing the constraint using a multiplicative constant puts more weight on the values of $\mathbf{k}^*(\mathbf{u})$ when $J^*(\mathbf{u})$ is small.

The choice of the relaxation constant $\alpha$ can be made to ensure the existence of a parameter that is ``acceptable'' with a certain probability.
For instance, given that $J>0$, $\Gamma_{\alpha}(\mathbf{k})$ is increasing with respect to $\alpha$ for any $\mathbf{k}\in\Kspace$. We can then focus on the smallest value of $\alpha$ such that $\Gamma_\alpha$ reaches a certain level of confidence $p\in[0,1]$. This leads to the definition of $\checkap$
\begin{align}
  \checkap &= \inf\left\{ \alpha\geq 1 \mid \exists \checkkp \in \Kspace,\, \Gamma_{\alpha}(\checkkp) \geq p \right\} \nonumber \\
   &= \inf \left\{ \alpha \geq 1 \mid \max_{\mathbf{k}\in\Kspace} \Gamma_{\alpha}(\mathbf{k}) \geq p \right\}   \label{eq:def_alpha_check}
\end{align}
that is the smallest $\alpha$, such that there exists a particular $\checkkp \in \Kspace$ for which $J(\checkkp,\mathbf{U}) \leq \checkap J^*(\mathbf{U})$ with probability $p$.
$\{\mathbf{k}_p \text{ for } p \in [0;1] \}$ defines the relative-regret family of robust estimators (\RRE).
This formulation can be linked to the quantiles and Value-at-Risk of the random variable $\max_{\mathbf{k}} \{ \frac{J(\mathbf{k},\mathbf{U})}{J^*(\mathbf{U})}\}$, that is a measure of risk usually applied in the financial sector (see~\cite{rockafellar_deviation_2002}). For different levels $p$, and thus different $\alpha_p$, Figure~\ref{fig:illu_alpha_p} shows examples of $\Gamma_{\alpha_p}$ for $\JBHS$. By definition, the associated $\mathbf{k}_p$ is then the first value for which $\Gamma_{\alpha_p}$ reaches $p$. We can see that changing the level $p$ shifts the maximiser of $\Gamma_{\alpha_p}$, and that for small $\alpha$, $\Gamma_{\alpha}(\mathbf{k}_1)$ is also very small. This indicates that $\mathbf{k}_1$ is located quite far from the conditional minimisers, and arise as a compromise when the relaxation is large enough.

\begin{figure}[!ht]
\centering
\includegraphics[width = 0.7\textwidth]{Figures/illu_alpha_p.pdf}
\caption{Illustration of the influence of different levels $p$ on $\Gamma_{\alpha_p}$ and on $\mathbf{k}_p$ for $\JBHS$}
\label{fig:illu_alpha_p}
\end{figure}

@@

By considering the particular case where $p=1$, a relation between $\checkk_1$ and $\checka_1$ can be derived, by using Eq.~\eqref{eq:def_alpha_check} and the strict positivity of the cost function:
\begin{equation}
  \label{eq:def_alpha1}
  \Prob_{\mathbf{U}}\left[\frac{J(\mathbf{k}_1,\mathbf{U})}{ J^*(\mathbf{U}) } \leq \checka_1\right]=1
\end{equation}
it follows then that 
\begin{equation}
  \label{eq:upper_bound_alpha_check}
  \checka_1 = \sup_{\mathbf{u}\in\Uspace} \frac{J(\mathbf{k}_1,\mathbf{U}=\mathbf{u})}{J^*(\mathbf{u})} =  \inf_{\mathbf{k}\in\Kspace} \left\{ \sup_{\mathbf{u}\in\Uspace} \frac{J(\mathbf{k},\mathbf{U}=\mathbf{u})}{J^*(\mathbf{u})} \right\} 
\end{equation}

This alternative definition of $\checka_1$ is useful to estimate it quickly whenever a grid over $\mathbb{K}\times\mathbb{U}$ has already been evaluated, thus avoiding multiple computations of $\Gamma_{\alpha}(\mathbf{k})$ until its maximum reaches $1$. From Eq.~\eqref{eq:upper_bound_alpha_check}, choosing a level of confidence equal to $1$ is then equivalent to looking for the worst-case scenario of the ratio $\frac{J(\mathbf{k},\mathbf{U}=\mathbf{u})}{J^*(\mathbf{u})}$. Therefore it may suffer from the same pitfall as the worst-case approach. As mentioned in the introduction, this is not suited for random variable with unbounded support as it may return over-conservative solution, provided that a solution exists in the first place.
